---
title: "Gravity Model Part 1"
author: "Ivanna Tindle"
date: "`r Sys.Date()`"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(readxl)
library(tidyverse)
library(ggplot2)
library(rlist)
library(rmarkdown)

#Set-up
```

## Coding assignment: Gravity Model

## **Data Retrieval:**

**Variables:**

$X_{ij}$ - exports between country i and country j for all i and j.

$GDP_i$ - real GDP in purchasing power equivalent \$PPE dollars for all i.

$d_{ij}$ - a distance proxy between countries i and j for all i and j.

**Data:**

GDP: US \$, constant prices, constant PPPs, reference year 2015, for 2021

Link: <https://stats.oecd.org//Index.aspx?QueryId=126176>

Exports -- Trade in value by partner countries (Edition 2022)

Link: <https://www.oecd-ilibrary.org/trade/data/oecd-quarterly-international-trade-statistics/trade-in-value-by-partner-countries-edition-2022_3928be98-en>

Distance -- Distance between country i's and j's capital using GeoDist database.

Link: <http://www.cepii.fr/CEPII/en/bdd_modele/bdd_modele_item.asp?id=6>

All files downloaded to same folder.

## Load Data

```{r}
# Change working directory as needed.
setwd("~/R/Gravity")

#May need to change file names here due to saved file name

GDP <- read_csv("SNA_TABLE_GDP_2021.csv", 
    col_types = cols(Year = col_number(), 
        Value = col_number()))

export <- read_csv("QITS-2022-1-EN-20230503T012136.csv", show_col_types = FALSE)

dist_ij <- read_excel("dist_cepii.xls")

```

Data Cleaning

```{r}
# Removing unnecessary variables

GDP <- GDP %>% select(LOCATION,Value) %>%
  rename(GDP = Value) 

export <- export %>% 
  filter(FLOW=="EXP") %>%
  filter(TIME==2021) %>%
  select(LOCATION, PARTNER, Value) %>% 
  rename(Exports = Value)


dist_ij <- dist_ij %>% 
  select(iso_o, iso_d, dist) %>%
  rename(LOCATION = iso_o,
         PARTNER = iso_d)

# Creating a copy of GDP for merging. 
GDP2 <- GDP %>% 
  rename(
  PARTNER=LOCATION,
  GDPj=GDP)

```

## Creating Data Set

```{r}
Data <- merge(dist_ij,export,by=c("LOCATION","PARTNER")) %>% 
  merge(GDP, by="LOCATION") %>% 
  rename(GDPi = GDP) %>% 
  merge(GDP2, by= "PARTNER") %>% 
  relocate(Country_i = LOCATION) %>% 
  relocate(GDPi, .after = Country_i) %>% 
  relocate(Country_j = PARTNER, .after = GDPi) %>% 
  relocate(GDPj, .after = Country_j)

Data <- Data %>% arrange(Country_i)

head(Data)

```

## Reshaping Data Set

```{r}
# In my best understanding, pivot_wider or pivot_longer is the equivalent to Matlab's reshape command. 

Data_wide <- Data %>%
 pivot_wider(
  names_from = Country_j,
  values_from = c(GDPj,dist,Exports),
  names_sep = "_"
)
head(Data_wide)
summary(Data_wide)
```
